---
title: "PA1_template"
author: "Simeon"
date: "Sunday, May 17, 2015"
output: html_document
---


Downloading and Loading Data

```{r, echo=FALSE}
setwd("C:/Users/Simeon/Desktop/R Language")
if(!file.exists("./Reproducible")){dir.create("./Reproducible")}
filename<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(filename, destfile = "./Reproducible/Dataset.zip")
unzip(zipfile = "./Reproducible/Dataset.zip", exdir = "./Reproducible")
setwd("C:/Users/Simeon/Desktop/R Language/Reproducible")
dataset <- read.csv("activity.csv")
dataset$date <- as.Date(dataset$date, "%Y-%m-%d")
```

Transforming dataset into suitable format

```{r, echo=FALSE}
dataset$date <- as.Date(dataset$date, "%Y-%m-%d")
```

Calculate the mean number of steps taken per day

```{r, echo =FALSE}
DailySteps <- aggregate(steps ~ date, data = dataset, sum, na.rm=TRUE)
hist(col = "green", DailySteps$steps, main = "Total Steps Taken per Day", xlab = "Total Steps")
```

Mean number of steps taken per day

```{r, echo=FALSE}
MeanSteps <- mean(DailySteps$steps)
```

The mean number of steps per day is 'r MeanSteps'.

Median number of steps taken per day

```{r, echo=FALSE}
MedianSteps <- median(DailySteps$steps)
```

The median number of steps per day is 'r MedianSteps'.

What is the average daily activity pattern?

```{r, echo=FALSE}
IntervalSteps <- aggregate(steps ~ interval, data = dataset, mean, na.rm = TRUE)
plot(steps ~ interval, data = IntervalSteps, type = "l", main = "Average number of steps across Intervals", xlab = "Intervals", ylab = "Average no. of Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=FALSE}
maxsteps <- IntervalSteps[which.max(IntervalSteps$steps), "interval"]
```

The interval with maximum number of steps is 'r maxsteps'.



Total number of rows with "NA":

```{r, echo=FALSE}
length(which(is.na(dataset$steps)))
```

Using Interval Mean to fill in missing values
```{r}
DuplicateData <- dataset
for (i in 1:nrow(DuplicateData)) {
      if (is.na(DuplicateData$steps[i])) {
            index <- DuplicateData$interval[i]
            value <- subset(IntervalSteps, interval==index)
            DuplicateData$steps[i] <- value$steps
      }
}
```

Number of steps taken each day (After filling in missing values)
```{r, echo=FALSE}
DailyAverageSteps <- aggregate(steps ~ date, data = DuplicateData, sum)
hist(DailyAverageSteps$steps, col = "blue", xlab = "Total Number of Steps", 
     ylab = "Frequency", main = "Histogram of Total Number of Steps taken each Day")
```

Mean of Total Steps taken per day

```{r, echo=FALSE}
mean(DailyAverageSteps$steps)
```

Median of Total Steps taken per day

```{r, echo=FALSE}
median(DailyAverageSteps$steps)
```

Do these values differ from the estimates from the first part of the assignment?

Only the median value differs.

What is the impact of imputing missing data on the estimates of the total daily number of steps?

Since the mean for each interval was input into the missing data, the mean stays the same.



Are there differences in activity patterns between weekdays and weekends?

```{r, echo=FALSE}
DuplicateData$date <- as.Date(DuplicateData$date, "%Y-%m-%d")
DuplicateData$day <- ifelse(weekdays(DuplicateData$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
DaySegregation <- aggregate(steps ~ interval + day, data = DuplicateData, mean)
library(lattice)
xyplot(steps ~ interval | factor(day), data = DaySegregation, type="l", aspect = 1/2)
```
